Approximation as Differentiation

### 2025-06-09 main/first-order.tex
Much pared back FinVect.

The category $\FinVect$ of finite vector spaces over $\RR$ (with addition of real numbers written $a + b$ and
multiplication written $a \mult b$) is semi-additive.

\begin{definition}[Finite dimensional vector space over $\RR$]
Up to isomorphism, an \emph{$n$-dimensional vector space over $\RR$} is the set $\RR^n$ equipped with
component-wise addition of vectors and scalar multiplication (also written $+$ and $\mult$), where $(\RR^n,+)$
is an abelian group (with identity $0$ and additive inverse $-v$ again defined component-wise), and where the
vector operations are compatible with the field operations in that the following equations hold \todo{probably
more detail than we need to go into}:

\vspace{-4mm}
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
\begin{align*}
   1 \mult v &= v \\
   (a \mult b) \mult v &= a \mult (b \mult v)
\end{align*}
\end{center}
\end{minipage}%
\begin{minipage}[t]{0.45\textwidth}
\begin{center}
\begin{align*}
   (a + b) \mult v &= (a \mult v) + (b \mult v) \\
   a \mult (u + v) &= (a \mult u) + (b \mult v)
\end{align*}
\end{center}
\end{minipage}
\end{definition}

\vspace{1mm}
\noindent Composition is with bilinear. The direct sum $V \biprod W \iso \RR^{m + n}$ given by the Cartesian
product is a biproduct $(V \biprod W, \biinj_{V}, \biinj_{W}, \pi_1, \pi_2)$ with $\biinj_{V} =
\prodM{\id_{V}}{0_{V,W}}$ and $\biinj_{W} = \prodM{0_{V,W}}{\id_{W}}$.

\begin{definition}
Define $\FinVect$ to be the category which has as objects $\RR^n$ all finite-dimensional vector spaces, and as
morphisms $f: \RR^m \to \RR^n$ all functions $f$ satisfying $f(u + v) = f(u) + f(v)$ and $f(a \mult v) = a
\mult f(v)$, with $f$ preserving the abelian group structure automatically.
\end{definition}

$\FinVect$ is enriched in $\Ab$, the category of abelian groups \todo{Introduce $\CMon$-enrichment first, then
perhaps $\Ab$-enrichment parenthetically}. For any objects $V = \RR^m$ and $W = \RR^n$ the hom-set
$\FinVect(V,W)$ is an abelian group, with identity $0_{V,W}$ given by the constant map $v \mapsto 0_W$ and
addition and additive inverse given pointwise:

\vspace{-4mm}
\begin{center}
\begin{minipage}[t]{0.35\textwidth}
\begin{align*}
(f + g)(v) &= f(v) + g(v)
\end{align*}
\end{minipage}%
\begin{minipage}[t]{0.35\textwidth}
\begin{align*}
(-f)(v) &= -f(v)
\end{align*}
\end{minipage}
\end{center}

>>>>

The category $\FinVect$ of finite vector spaces over $\RR$ is semi-additive.

\begin{definition}
Define $\FinVect$ to be the category which has as objects $\RR^n$ all finite-dimensional vector spaces
$(\RR^n, +, \mult)$, and as morphisms $f: \RR^m \to \RR^n$ all functions $f$ satisfying $f(u + v) = f(u) +
f(v)$ and $f(a \mult v) = a \mult f(v)$.
% with $f$ preserving the zero and additive inverse structure automatically.
\end{definition}

%  with addition of real numbers written $a + b$ and multiplication written $a \mult b$

% \begin{definition}[Finite dimensional vector space over $\RR$]
% Up to isomorphism, an \emph{$n$-dimensional vector space over $\RR$} is the set $\RR^n$ equipped with
% component-wise addition of vectors and scalar multiplication (also written $+$ and $\mult$), where $(\RR^n,+)$
% is an abelian group (with identity $0$ and additive inverse $-v$ again defined component-wise), and where the
% vector operations are compatible with the field operations in that the following equations hold:
%
% \vspace{-4mm}
% \begin{minipage}[t]{0.45\textwidth}
% \begin{center}
% \begin{align*}
%    1 \mult v &= v \\
%    (a \mult b) \mult v &= a \mult (b \mult v)
% \end{align*}
% \end{center}
% \end{minipage}%
% \begin{minipage}[t]{0.45\textwidth}
% \begin{center}
% \begin{align*}
%    (a + b) \mult v &= (a \mult v) + (b \mult v) \\
%    a \mult (u + v) &= (a \mult u) + (b \mult v)
% \end{align*}
% \end{center}
% \end{minipage}
% \end{definition}

$\FinVect$ is enriched in $\CMon$ (in fact in $\Ab$, the category of abelian groups, but $\CMon$ will do for
our purposes). For any objects $V = \RR^m$ and $W = \RR^n$, the hom-set $\FinVect(V,W)$ provides pointwise
addition of morphisms $(f + g)(v) = f(v) + g(v)$; the constant map $v \mapsto 0$ is a zero morphism, and
composition is bilinear. Moreover the direct sum $V \biprod W \iso \RR^{m + n}$ given by the Cartesian product
is a biproduct, with $\biinj_{V} = \prodM{\id}{0}$ and $\biinj_{W} = \prodM{0}{\id}$.
